---
title: "Comparison of the Exponential Distribution with the Central Limit Theoreom"
author: "Gene Kaufman"
date: "February 7, 2016"
output: pdf_document
---

#Overview
The Exponential Distribution can be simulated with the R function *rexp(n, lambda)*, where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Using lambda = 0.2, I will investigate the distribution of 1000 simulations of size 40 exponentials and compare to the Central Limit Theorem.

#Simulations 
First, load some libraries and set some options
```{r setoptions, message=FALSE}
require(knitr)
opts_chunk$set(echo=TRUE, results="asis", warning=FALSE, message=FALSE)
```

Initialize Variables
```{r, init.vars,cache=TRUE}
lambda <- 0.2
n <- 40
num_sims <- 1000
```

Create Simulations
```{r, run.simul.1,cache=TRUE}
set.seed(42) # Reproducibility!
```

Build matrix of `r num_sims` simulations of `r n` Exponentials
```{r, run.simul.2,cache=TRUE}
exp_dist <- matrix(rexp(n * num_sims,lambda),num_sims)
```

Calculate means of the simulations
```{r, run.simul.3,cache=TRUE}
exp_means <- apply(exp_dist, MARGIN=1,FUN=mean)
```

#Sample Mean versus Theoretical Mean: 
The Theoretical Mean of the Exponential Distribution is *1/lambda*
```{r compare1}
theory_mean <- 1/lambda
```

The average Sample Mean:
```{r compare2}
avg_sample_mean <- mean(exp_means)
```

Theoretical mean: `r theory_mean`, Sample mean: `r avg_sample_mean`

Plotting everything onto one chart makes it easy to compare that the Sample Mean and Theoretical Mean are pretty close:
```{r chart1, echo=FALSE}
theory_sd <- 1/lambda
numBreaks <- 100
# plot the averages as density probabilities
hist(exp_means
     ,probability=TRUE
     ,breaks = numBreaks
     ,xlab="Means"
     ,main="Exponential Distribution:\nSimulated Sample Means vs Theoretical Means"
     )

# add density plot of simulated sample means
lines(density(exp_means),lty=1)

# restructure theoretical mean to create a density plot on
# the same scale as the simulated sample means
xfit <- seq(min(exp_means), max(exp_means)
            , length=numBreaks
            )
yfit <- dnorm(xfit, mean=theory_mean, sd=(theory_sd/sqrt(n)))
lines(xfit, yfit
      , lty=2
      )

# add a vertical line at the theoretical mean
abline(v=theory_mean, lwd=3)

# add legend
legend('topright'
       , c("Theoretical","Simulated Samples")
       , lty=c(2,1)
       , cex=0.6
       , bty="n"
       )
```


#Sample Variance versus Theoretical Variance: 
The Theoretical Variance of the Exponential Distribution is the square of the Theoretical SD: *(1/lambda)^2* or *theory_sd^2*  
```{r compare2_1}
theory_sd <- 1/lambda
theory_var <- theory_sd ^ 2

```
However, we are comparing to a sample, so we have to account for the sample size:
```{r compare2_1a}
theory_sd_samp <- theory_sd / sqrt(n)
theory_var_samp <- theory_sd_samp ^ 2
```

The average Sample Variance:
```{r compare2_2}
avg_sample_var_means <- var(exp_means)
```

Theoretical Variance: `r theory_var_samp`, 

Average of Sample Variance Means: `r avg_sample_var_means`



#Distribution: 
Explain how the distribution is approximately normal. The Central Limit Theoreom states that the mean of large number of iterations of a distribution will be approximately normally distributed. Here we can see that the histogram of 
means has a very Gaussian look, centered around the distribution mean

```{r chart3}
hist(exp_means,
     main=paste("Means from Exponential Distribution:\nSimulations:", num_sims , "\nSample Size:", n ,"\nRate:", lambda),
     xlab="Means",
     breaks=20)
# add a vertical line at the theoretical mean
abline(v=theory_mean, lwd=3)

```
